\documentclass[a4,12pt]{amsart}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[american]{babel}
\usepackage{amssymb,amsbsy,amsmath,amsfonts,amssymb,amscd}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage{eurosym}
\usepackage[left=2.5cm, right=2.5cm, top=2.5cm,bottom=2cm]{geometry}

\newcommand{\sage}{\href{http://www.sagemath.org/}{\texttt{Sage}}\xspace}
\newcommand{\sagecombinat}{\href{http://wiki.sagemath.org/combinat/}{\texttt{Sage-Combinat}}\xspace}
\usepackage{color}
\newcommand{\TODO}[2][To do: ]{{\textcolor{red}{\textbf{#1#2}}}}
%\newcommand{\TODO}[2][To do: ]{}

\usepackage{enumitem}
\setitemize{leftmargin=4ex, noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% A wider quote
\renewenvironment{quote}{%
  \list{}{%
    \leftmargin0.5cm   % this is the adjusting screw
    \rightmargin\leftmargin
  }
  \item\relax
}
{\endlist}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[SageMath in Europe]{SageMath in Europe: Mutualized development\\
  of computer exploration environment\\
  for research in combinatorics, algebra, and beyond
}
% Titre français:
% Développent logiciel mutualisé pour la recherche en combinatoire et au delà
\date{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\ \vspace{-1.5cm}
\hrule
\medskip
\maketitle
\ \vspace{-1cm}
\hrule
%\subsection*{Long title}

%\TODO{From the application guide: maximum 150 caractères}

%\textbf{Mutualized software development\\for research in combinatorics and beyond}

%\textbf{Sage-Combinat: fostering code sharing in combinatorics and beyond, for computer exploration and research}
% Keywords: in the French community?
% Research on algorithms, benchmarking, ...
% Sage-Combinat: un peu restrictif
% Mission: “To improve the open source mathematical system \sage as an extensible toolbox for computer exploration in combinatorics and beyond, and foster code sharing between researchers in this area”.


\section{Context and scientific theme}
\TODO{Mettre en avant le fait que Sage a des bugs critiques et qu'on a besoin de travail technique pour en venir a bout.}

% \TODO{From the application guide: 2 à 3 pages maximum Le contenu de
%   cette section permet une appréciation selon le premier critère
%   d’évaluation (Intérêt des objectifs scientifiques et
%   technologiques).  Donner les informations permettant d’évaluer
%   l’intérêt des objectifs scientifiques et technologiques, décrire
%   brièvement la méthodologie et/ou la capacité à générer des
%   résultats, le potentiel d’avancée dans le domaine, l’ambition, la
%   nouveauté, le potentiel de rupture.}

Since the 50's computer exploration has been an instrumental tool for
research in mathematics; in some cases, computers have even used for
proving results like for the famous four color theorem, or the
alternating sign matrices conjecture.
%\TODO{the analogue of telescopes for astrophysicists}.
To support this, much efforts have been spent on the software
development. As the sophistication of the required computations
increased, supported by the boom of the available computational power,
it was progressively realized that single topic/single person software
projects, or even projects at the level of a single research team
simply did not scale. It was essential to share the software
development at the level of a research community, both in term of
breath of features and mutualization of development efforts. This lead
to the apparition in the 90's of many specialized open source
packages, like GAP for group theory, Maxima for symbolic computation, or
Pari/GP for number theory.

Yet, this proved insufficient in many areas like combinatorics, as
computer exploration often requires simultaneously a wide variety of
objects and features which only a general purpose software system can
provide and support. At the same time, the then existing general
purpose systems like \texttt{Maple} or \texttt{Mathematica} prevented
the scaling of existing packages because they lacked basic computer
science paradigms; even more important, they were not open source.

\subsection*{Originality of the Sage project}

To tackle this situation, William Stein started in 2005 the \textbf{\sage
project, a free open-source mathematics software system}. This project
soon crystallized a large international community of contributors; now
more than 300 researchers and teachers in maths and computer
science. It already played a fundamental role in hundreds of
scientific publications, and its usage for \texttt{higher education}
is spreading.

Building on the very active Python and Scientific Python stack
combined with the power of dozens of open-source mathematical packages
(GAP, Singular, or FLINT to name but a few), \sage provides a full
featured research environment including a rich notebook interface for
interactive computations and visualization, as well as support for
intensive batch computations, for example on the cloud. \sage is also
the computational workhorse behind sister database projects like
\textsc{LMFDB}, the online database of L-functions, modular forms, and
related objects.

Thanks to a high level programming framework, most of the code can be
written generically and reused in widely different contexts, including
\textbf{physics and computer science}. The induced cross-discipline
interactions both between developers and with users
are a regular source of cross fertilization, and already lead to
several joint research projects.

\sage \textbf{is particularly beneficial for young researchers},
giving them access to a platform with a wide array of features, and
allowing them to focus on those computational aspects directly related
to their own research topic. With, at the end, the possibility and
incent to contribute back their developments to the community
\emph{and get credit for it}.

Collaborative software development involves lots of interactions
within the community, around software and user interface design,
algorithmic, code review, governance, and maintenance. Those
interactions are organized using modern collaborative tools for open
source agile development (distributed version control, mailing lists,
ticket server, testing tools, ...), and following best practices. A
strong emphasis is put on training, in particular at the occasion of
numerous week-long workshops like the upcoming \emph{Sage Days 60} in
India.

\TODO{B) Les challenges et l'originalité de la solution}

\subsection*{Scaling further toward 2020}

After 10 years of existence, \textbf{\sage has proven its viability
  and leadership}, surviving the many hurdles faced by new open source
software projects, reaching a comfortable community size, and proving
the adequacy of its ``developed by users, for users'' model.  Yet
\sage has not yet been systematically adopted, and still \textbf{much
  good research time is wasted developing and redeveloping tools by
  lack of sharing and coordination}.

By design, the development of \sage has been consciously designed to
run on a small budget. This was a key feature toward its independence
and its early wide adoption across research teams and international
frontiers. In particular, it is ``developed by users, for users'':
most if not all of the development is achieved by the researchers
themselves and justified w.r.t. their funding institutions by the
immediate application to their own research projects. Core development
is taken care of by permanent researchers that can afford to invest
time over the long run (10 years). Coordination is achieved
electronically using modern collaborative development tools
(distributed version control, ticket server, mailing lists, ...).

So far, \sage has been running viabily on a small budget for its size (300
developers), thanks to its ``developed by users for users'' model. Yet,
long term non mathematical features have been lagging behind because they cannot be
split in pieces small enough to be implemented as a side product of research
projects: portability (including native Windows support), build infrastructure
and modularization, packaging and distribution, improved user interfaces, core
support for parallelism. Those features are critical to let \sage scale to
mainstream and assume its leadership.
\TODO{FINISH THIS SENTENCE}
\textbf{For those, the
development model needs to be backed up by full time research
ingeneers}.

It is also vital to organize regular physical meetings (missions,
invitations, workshops) for fast coordination; this becomes even more
critical when it comes to train massively new developers and users.
Finally, productive software development requires powerful hardware for
compilation, regression testing, continuous integration,
benchmarking.
%
Up to now, this has been funded as an aside of research
projects (\TODO{cite them}), or through the case by case generosity of
labs, taken on their recurrent budget (e.g. developers meeting and
missions funded by the LITIS in Rouen, the LRI and LMO in Orsay,
... \TODO{expand to the European scale}), but it has reached its
limits, with much time wasted on patching together little pieces of
funding.
%



\textbf{To scale further, \sage needs direct funding in addition to
  indirect funding through research projects.} Such funding already
exists in the United States, e.g. with large (500k\$) NSF
Computational Mathematics grants; the boosting effect of those on the
community is already shining there (workshops, ICERM semester, many
new recruits, ...).

\medskip
\textbf{The purpose of this proposal is to trigger a similar boost in
  Europe.}

\section{Outreach}

\TODO{C) Les retombées scientifiques et sociétales}

\TODO{Insert this somewhere: Dissemination: Sharing code increases our
  visibility, spreads our research results to other scientific areas,
  and benefits education.}

\subsection{Teaching}

We plan to integrate \sage into the curriculum of the courses at our
universities. Many undergraduate and graduate courses in number theory
and combinatorics already involve computational components. Moreover,
since this year, Sage plays an important role in the Agrégation. With
the dwindling resources at the universities, a free and open-source
alternative to the currently used systems % such as {\sc Mathematica}
%and {\sc MatLab}
is most welcome. Since \sage is based on Python, one of the top five
programming languages, this also gives the students valuable
education.  Inspired students can furthermore look at the algorithms
or even contribute back to \sage.

\TODO{Reuse stuff from the NSF grant}

\section{Consortium}

\TODO{Expand}

The consortium will tentatively have participants in Austria (Linz:
CS), France (Rouen: CS, Bordeaux: CS, Lyon: math, Marseille: math,
Marne: CS, Paris: math, CS, physics, Saclay: CS), Germany (Berlin,
Jena: math), Switzerland (Zürich: math), Great Britain (Bristol:
math), Ireland (Dublin: physics), Quebec (Montréal, joint lab
CRM/CNRS: math).

The PI's have developed a strong community building experience which
they gained in particular along the \textsc{MuPAD-Combinat} project
(now \sagecombinat) that progressively gathered a community of 42+
contributers since it's foundation in 2000.

On the French side, the consortium will be built around a core team
that has been closely collaborating around Sage in France and Québec
for a long time.
%Delecroix, Cohen, Labbé, Lelièvre, Hivert, Monteil, Mezzarobba,
%Saliola, Thiéry.
In particular they coorganized and participated to a stream of events
that included (international) developer meetings, training sessions,
and development of educational material.
%
This core team covers a wide continuum of research themes in discrete
mathematics, dynamical systems, algebra, or number theory, with deep
connections to neighbor fields like symbolic computations or
statistical physics. It has developed an international network of
collaborators that will be involved in the project, including a large
number of PhD students and postdocs.

\TODO{Expand}

\input{actions}

% \TODO{From the application guide: (1 à 2 pages maximum) Le contenu de
%   cette section permet une appréciation selon le troisième critère
%   d’évaluation (Cohérence de la pré-proposition par rapport aux
%   objectifs du projet) Donner les informations relatives aux
%   compétences requises pour mener le projet objet de la
%   pré-proposition, en précisant l’identité du ou des scientifique(s)
%   impliqué(s), l’identification des Partenaires 4 auxquels il(s)
%   est(sont) rattaché(s) et 2-3 références pertinentes dans le domaine
%   en lien direct avec la pré-proposition (publications, faits
%   marquants de R\&D, brevets, prix scientifiques, produits, procédés,
%   licences, services...), et tout autre élément permettant de juger de
%   la qualité des déposants et du consortium le cas échéant.}

% The consortium is built around a core team that has been closely
% collaborating around Sage in France and Québec for a long time:
% Delecroix, Cohen, Labbé, Lelièvre, Hivert, Monteil, Mezzarobba,
% Saliola, Thiéry. In particular they coorganized and participated to a
% stream of events that included (international) developer meetings,
% training sessions, and development of educational material:
% \begin{itemize}
% \item \href{http://wiki.sagemath.org/GroupeUtilisateursParis}{Groupe
%     d'utilisateurs Sage en région parisienne}: a monthly meeting;
%   %Labbé, Hivert, Thiéry, Valibouze, et al.
% \item \href{http://www.ragaad.org/bobo2012/}{\emph{CIMPA discrete math research school}},
%     % : «Mathématiques discrètes: aspects combinatoires, dynamiques et algorithmiques»}},
%     Bobo-Dioulasso (Burkina Faso), Fall 2012;
%     %Lelièvre, Delecroix, Monteil, Thiéry et al.
% %\item \href{http://smai.emath.fr/smai2011/programme_slides.php}{\emph{Congrès SMAI}}, Mini-symposium \sage, Guidel May 2011;
% %  Mezzarobba, Thiéry et al.
% \item Sage course, \emph{\'Ecole Jeunes Chercheurs Informatique
%     Math\'ematique}, %April
%   Perpignan, 2013;
%   % : Monteil, Cohen,  Delecroix, Lelièvre.
% \item Three week-long developer meetings in Cernay, Orsay (2009, 2010, 2012);
%   % Borie, Cohen, Boussicault, Delecroix, Hivert, Labbé, Monteil, Saliola, Thiéry et al.
% \item \href{http://wiki.sagemath.org/combinat/FPSAC13}{Sage Days 49},  Orsay, %June
%   2013;
%   % , Cohen, Hivert, Mezzarobba, Monteil, Thiéry.
% %\item
%   \href{http://wiki.sagemath.org/days38}{Sage Days 38}, Montréal, %May
%   2012;
%   % , Brlek, Hivert, Labbé, Saliola, Thiéry et al.
% %\item
%   \href{http://wiki.sagemath.org/days28}{Sage Days 28}, %: dynamics, geometry, combinatorics},
%   Orsay, %January
%   2011;
%   %, Borie, Cohen, Delecroix, Hivert, Labbé, Lelièvre, Monteil, Thiéry et al.
% \item \href{https://www.lirmm.fr/arith/wiki/MathInfo2010/SageDays}{\emph{Sage Days 20}},
%   during the \emph{Math-Info 2010} thematic month at CIRM, Luminy;%, February 2010
%   % : a one week workshop during the \emph{Math-Info 2010} thematic month at
%   %CIRM, Luminy, France, February 2010; Delecroix, Hivert, Labbé,
%   %Monteil, Saliola, Thiéry et al.
% \item Introductory book «\emph{Calcul Mathématique avec Sage}»,
%   % Cohen, Thiéry, Mezzarobba contributed a third of the chapters.
% %\item
%  Stream of Sage thematic tutorials.
%   %: Cohen, Labbé, Lelièvre, Hivert, Saliola, Thiéry et al.
% \end{itemize}


%\bibliographystyle{alpha}
%\bibliography{project-description}
\end{document}
